Structuring Theories with Implicit Morphisms

### 2018-09-27 mmt.tex
update

  axiom_assoc		: ⊦ ∀[x]∀[y]∀[z] x ∘ (y ∘ z) ≐ (x ∘ y) ∘ z ❙
  axiom_right_unit	: ⊦ ∀[x] x ∘ e ≐ x ❙
  axiom_left_unit	: ⊦ ∀[x] e ∘ x ≐ x ❙
  axiom_inverse : ⊦ ∀[x] x ∘ (x⁻¹) ≐ e ❙
❚
\end{mmtcode}
\begin{mmtcode}
theory DivisionGroup : ?Meta =
  universe : type ❘ # U prec 1 ❙
  div : U ⟶ U ⟶ U ❘ # 1 / 2 ❙
  nonEmpty : U ❘ # e prec 1 ❙
  // axioms ... ❙
❚
\end{mmtcode}
where $\GS,\US,\RS$ are delimiters and the (optional) \lstinline|# <not>| is an optional notation. \lstinline|?Meta| is an intended meta-theory containing e.g. the symbols $\forall,\doteq$ etc.
\end{example}

@@

  inverse : U ⟶ U ❘  # 1 ⁻¹ ❙
  axiom_inverse : ⊦ ∀[x] x ∘ (x⁻¹) ≐ e ❙
❚
\end{mmtcode}
The flat variant $\cn{Group}^\flat$ we get by making all included symbols (from $\cn{Monoid}$) available to $\cn{Group}$, resulting in exactly the theory as presented earlier in \autoref{ex:thgroup}.
\end{example}

@@

Also note, that because $\flt{S}$ is a \textit{set} of declarations, the include relation is transitive: if $t$ includes $s$ via two different paths, $\flt{t}$ only contains one copy of the declarations of $\flt{s}$.
The situation is slightly more complicated for morphisms: if a morphism out of $t$ includes two different morphisms out of $s$, these have to agree.
Therefore, flattening and well-formedness must be distinguished:
  \[\fltd{\sigma}{\icl{M}} = \flt{M}\]
under the condition $\flt{M}$ agrees with $\flt{\sigma}$ on any constant that is in the domain of both.
\end{example}

>>>>

  axiom_assoc		: ⊦ ∀[x]∀[y]∀[z] x ∘ (y ∘ z) ≐ (x ∘ y) ∘ z ❙
  axiom_right_unit	: ⊦ ∀[x] x ∘ e ≐ x ❙
  axiom_left_unit	: ⊦ ∀[x] e ∘ x ≐ x ❙
  axiom_inverse : ⊦ ∀[x] x ∘ (x⁻¹) ≐ e ❙
❚
\end{mmtcode}
\begin{mmtcode}
theory DivisionGroup : ?Meta =
  universe : type ❘ # U prec 1 ❙
  div : U ⟶ U ⟶ U ❘ # 1 / 2 ❙
  nonEmpty : U ❘ # e prec 1 ❙
  // axioms ... ❙
❚
\end{mmtcode}
where $\GS,\US,\RS$ are delimiters and the \lstinline|# <not>| components are (optional) notations. \lstinline|?Meta| is an intended meta-theory containing e.g. the symbols $\forall,\doteq$ etc.
\end{example}

@@

  inverse : U ⟶ U ❘  # 1 ⁻¹ ❙
  axiom_inverse : ⊦ ∀[x] x ∘ (x⁻¹) ≐ e ❙
❚
\end{mmtcode}
\end{example}

@@

Also note, that because $\flt{S}$ is a \textit{set} of declarations, the include relation is transitive: if $t$ includes $s$ via two different paths, $\flt{t}$ only contains one copy of the declarations of $\flt{s}$.
The situation is slightly more complicated for morphisms: if a morphism out of $t$ includes two different morphisms out of $s$, these have to agree.
Therefore, flattening and well-formedness must be distinguished:
  \[\fltd{\sigma}{\icl{M}} = \flt{M}\]
under the condition $\flt{M}$ agrees with $\flt{\sigma}$ on any constant that is in the domain of both.

The flat variant $\cn{Group}^\flat$ of our refactored theory $\cn{Group}$ from \autoref{syn:incl} is hence given by copying all included symbols (from $\cn{Monoid}$) over to $\cn{Group}$, resulting in exactly the theory as presented earlier in \autoref{ex:thgroup} except for global identifiers.
\end{example}
